### 2023-03-06 paper/paper.Rmd
started intro paper

---
title: |
  Introducing STRAIN: A simple and interpretable pass rush metric inspired by materials science
short_title: STRAIN
authors: 
- name: "Author 1"
  affiliation: "Address"
  city: "City"
  country: "Country"
  email: "email1\\@example.com"
- name: "Author 2"
  affiliation: "Address2"
  city: "City2"
  country: "Country2"
  email: "email2\\@example.com"
abstract: |
  Abstract goes here
keywords: tracking data, Bornn legacy
bibliography: "bibliography"
output: rticles::rss_article
---

# Introduction

A paragraph on tracking data

<!--
TODO: update citations after they're published
-->

See @Baumer2023Big, @Macdonald2020Recreating, and @Kovalchik2023Player for more details

Importance of linemen

Previous metrics: Greg's lit review, their flaws

Our contribution

# Data

# Methods

# Results

example play

positional curves

rankings (indiv, add more for team-level offense/defense)

corr with pressure rate = (hits + sacks + hurries) per snap

meta metrics: discriminant, stability, independence

@Franks2016Meta