### 2023-03-06 paper/paper.Rmd
paper sketch + everything from Kaggle notebook

---
title: |
  Introducing STRAIN: A simple and interpretable pass rush metric inspired by materials science
short_title: STRAIN
authors: 
- name: "Author 1"
  affiliation: "Address"
  city: "City"
  country: "Country"
  email: "email1\\@example.com"
- name: "Author 2"
  affiliation: "Address2"
  city: "City2"
  country: "Country2"
  email: "email2\\@example.com"
abstract: |
  Abstract goes here
keywords: tracking data, Bornn legacy
bibliography: "bibliography"
output: rticles::rss_article
---

**PAPER SKETCH**

1. Introduction

<!--
TODO: update citations after they're published
-->

Tracking data, big data bowl
See @Baumer2023Big, @Macdonald2020Recreating, and @Kovalchik2023Player for more details

Importance of linemen. Did all hard work. QB

Previous metrics: Greg's lit review, their flaws

Our contribution

2. Data

talk about data

3. Methods

4. Results

example play

positional curves

rankings (indiv, add more for team-level offense/defense)

corr with pressure rate = (hits + sacks + hurries) per snap

meta metrics: discriminant, stability, independence @Franks2016Meta

5. Discussion

# Introduction

<!--

Greg's long lit review

The theme of Big Data Bowl 2023 is about evaluating performance of linemen, both offensive and defensive, on passing plays. While the ideas presented here could be extended to offensive linemen, we have chosen to focus on measuring the performance of defensive linemen. Officially tracked by the NFL since 1982, the most well known statistic for evaluating defensive linemen on pass rushing plays is the sack, which occurs when a defensive player tackles the quarterback behind the line of scrimmage before the quarterback is able to throw a pass. Other counting statistics have been tracked by various entities such as hurries and hits as well as pressures, which is defined as the sum of hurries, hits, and sacks. More recently metrics such as Time In Pocket (TIP) and Pass Rusher Win Rate (PRWR) have also been proposed and are substantial improvements over the less sophisticated counting statistics. While PRWR is certainly a step in the right direction, it takes continuous data and converts it to a binary win/loss, which is generally not recommended in statistical analysis, thus, the metric we propose here is computed continuously over time just as the data is actually observed.

The metric we propose is based on the following simple question: What is considered a success for a pass rusher on a pass play? The simplest answer to this question, and the first defensive lineman statistics to be tracked by the NFL, is a sack. In terms of tracking data, one can think of a sack as reducing the distance between the pass rusher and the quarterback to 0. Therefore, a decent starting point for a continuous measure of pass rushing effectiveness shoould be based on how close the pass rusher is able to get to the quarterback with smaller distances indicating amore successful pass rush. However, a good measure of pass rushing effectiveness cannot simply be based on distance to the quarterback; Consider the following case: a pass rusher gets within two yards of the quarterback, but is then stopped and fails to get any closer for the rest of the play. This is not a successful rush because the pass rusher is not continuing to reduce the distance between themselves and the quarterback. Therefore, a good metric should consider the rate at which this distance is being reduced. However, just like distance, rate by itself is inadequate in assessing success of a pass rusher. Consider the following case: A defensive player is 50 yards away and moving towards the quarterback very rapidly. Their moving rate is essentially unimportant due to a significantly far distance away from the quarterback.

This means that in order to appropriately evaluate rusher performance, the distance from the quarterback and the rate at which this distance is being reduced must both be considered. Therefore, we propose STRAIN, a new measure for pass rusher effectiveness, which is computed as the ratio of the rate that the pass rusher is reducing the distance between themselve and the quarterback and distance between the pass rusher and the quarterback. In performing some research about this proposed quantity, we discovered that this concept actually already exists in materials science and is called the "strain rate" (which is where the name of our metric is derived). Given this connection, one can view the players as particles in some material and the defensive "particles" are attempting to exert pressure on the pocket in an attempt to compress and collapse this pocket around the quarterback. We think this analogy is just delightful.

What is considered a success for a pass-rusher on a pass play?
The simplest answer to this question is a sack.
In terms of tracking data, one can think of a sack as reducing the distance between the pass-rusher and quarterback to 0.
More generally, a good starting point for a continuous measure of pass-rushing effectiveness should be based on the **distance** between the rusher and quarterback, with smaller distances indicating greater success.
However, this naive approach has major drawbacks.
Consider the case where a rusher is within two yards of the quarterback, but is then stopped and fails to get any closer for the rest of the play.
This is not a successful rush because the pass-rusher is not continuing to reduce their distance to the quarterback.
Thus, another approach for pass-rushing assessment is to measure the **rate** at which the rusher is getting to the quarterback.
Yet, the rate by itself is also inadequate.
If a defender is 50 yards away and moving towards the quarterback, their moving rate is essentially unimportant due to a significantly far distance.

Therefore, in order to appropriately evaluate rusher performance, the **distance** from the quarterback *and* the **rate** at which this distance is being reduced must both be considered.
We propose **STRAIN**, a new measure for pass-rusher effectiveness, which is computed as the ratio of the aforementioned moving rate and distance quantities.
Our proposed metric has a nice interpretation and is inspired by the concept of "strain rate" in materials science.

-->

# Data

introduce data

# Methods

## Strain rate in materials science

In [*materials science*](https://en.wikipedia.org/wiki/Materials_science), [*strain*](https://en.wikipedia.org/wiki/Strain_(materials_science)) is the deformation of a material from stress.
The stretching of a rubber band is an example of strain: the longer the stretch, the larger the strain.
Formally, let $L(t)$ be the distance between two points of interest at time $t$, and $L_0$ be the initial distance between those two points.
Then the strain at time $t$ is defined as $\displaystyle s(t)={\frac  {L(t)-L_{0}}{L_{0}}}$.
Notice that this measure is unitless as a ratio of two quantities having the same dimension.
Furthermore, the [*strain rate*](https://en.wikipedia.org/wiki/Strain_rate) is the derivative of strain.
That is, ${\displaystyle {{s}'}(t)={\frac {ds}{dt}} = {\frac {v(t)}{L_{0}}}}$, where $v(t)$ is the rate at which the two points of interest are moving away from/towards each other.
Whereas strain is unitless, the strain rate is measured in inverse of time, usually $\text{seconds}^{-1}$.

## Application to pass-rushing in football

Motivated by its scientific definition, we draw an analogy between strain and pass-rushing in football.
**Just as strain rate is a measure of deformation in materials science, a pass-rusher's efforts involve the application of force/deformation against the offensive line, with the ultimate goal of breaking through the protection to reach the quarterback.
The players can be viewed as "particles" in some material and the defensive "particles" are attempting to exert pressure on the pocket with the aim of compressing and collapsing this pocket around the quarterback.**

In order to apply strain rate to measure NFL pass-rusher effectiveness, we make modifications to how this concept is traditionally defined.

Let $(x_{ijt}, y_{ijt})$ be the $(x, y)$ location on the field of player $j = 1, \cdots, J$ at frame $t = 1, \cdots, T_i$ for play $i = 1, \cdots, n$; and $(x^{QB}_{it}, y^{QB}_{it})$ be the $(x, y)$ location of the quarterback at frame $t$ during play $i$.

* The distance between player $j$ and the quarterback at frame $t$ during play $i$ is $f_{d_{ij}}(t) = \sqrt{(x_{ijt} - x^{QB}_{it})^2 + (y_{ijt} - y^{QB}_{it})^2}$.
* The rate at which player $j$ is moving towards the quarterback at frame $t$ during play $i$ is $\displaystyle f'_{d_{ij}}(t) = \frac{df_{d_{ij}}(t)}{dt}$.
* The STRAIN for player $j$ at frame $t$ during play $i$ is

$$\text{STRAIN}_{ij}(t) = \frac{-f'_{d_{ij}}(t)}{f_{d_{ij}}(t)} \,.$$

(*Note that to distinguish our metric from strain and strain rate in materials science, we write it in capital letters* (STRAIN) *for the remainder of this report.*)

Recall that by its materials science property, strain is generally defined to increase as the distance between two points **increases**.
In our football setting, the two points of interest are the pass-rusher and the quarterback, and we want our metric to increase as the distance between the pass-rusher and the quarterback **decreases**.
Thus, the negative sign in the numerator of our formula accounts for this. 
Additionally, rather than keeping the initial distance ($L_0$ as previously denoted) between two points constant over time, we update the initial position to be the player locations at the beginning of each frame.
This effectively gives us the STRAIN for each frame throughout a play.

Since we only observe these functions discretely in increments of 10 frames/second, an estimate for our proposed metric STRAIN is

$$\widehat{\text{STRAIN}}_{ij}(t) = \cfrac{-\cfrac{f_{d_{ij}}(t) - f_{d_{ij}}(t - 1)}{0.1}}{f_{d_{ij}}(t)} \,.$$

Notice that this quantity increases in two ways: 1) the rate at which the rusher is moving towards the quarterback increases, and 2) the distance between the rusher and the quarterback gets smaller.
Both of these are indications of pass-rusher effectiveness.
Finally, our statistic STRAIN is measured in inverse seconds, similar to strain rate.
(There is also an interesting interpretation of the reciprocal of our metric, namely, the amount of time required for the rusher to get to the quarterback at the current location and rate at any given time $t$.)

# Results

In the forthcoming analysis, we rely on the provided data of all passing plays from the first 8 weeks of the 2021 NFL season.
We include only the frames between the ball snap and when a pass forward or QB sack is recorded for each play.
We also remove all plays with multiple quarterbacks on the field, since we need a uniquely defined quarterback to calculate our measure.
Using PFF scouting data, we only consider defenders with role "Pass rush" for evaluation.

## Example play

To illustrate STRAIN for pass-rusher evaluation, we use a play from the 2021 NFL regular season week 6 matchup between the Las Vegas Raiders and Denver Broncos, which ended with [Broncos QB Teddy Bridgewater getting sacked by Raiders DE Maxx Crosby](https://www.raiders.com/video/de-maxx-crosby-sacks-qb-teddy-bridgewater-for-a-loss-of-6-yards-nfl).
The figure below consists of two elements:
at top, the locations of every Las Vegas (black) and Denver (orange) player on the field during the play, with Maxx Crosby highlighted in blue;
and at bottom, a line graph showing how Crosby's distance from the QB, velocity, and STRAIN change continuously throughout the play.
Note that the point size for each Las Vegas defender corresponds to the estimated STRAIN in each frame as the play progresses.

For the first 2.5 seconds, Crosby is being covered by a Denver player and unable to get close to the quarterback, hence the corresponding STRAIN values are virtually zero.
Suddenly, the STRAIN increases after Crosby is freed up and charges towards Bridgewater.
At 4 seconds after the snap, Crosby's STRAIN is 2.30, which means at his current moving rate, it will take Crosby about 0.43 (1/2.30) seconds to make the distance between him and the quarterback 0 (i.e. essentially sack the quarterback).
This is consistent with the final outcome of the play, as the sack takes place at the very last frame (4.4 seconds) where the estimated STRAIN for Crosby reaches its peak at 3.96.

## Positional STRAIN curves

The following plot displays the average STRAIN by position for the first 40 frames (4 seconds) after the snap.
We observe a clear difference in STRAIN between edge rushers (OLB's and DE's) and interior linemen (DT's and NT's).
Specifically, edge rushers have higher STRAIN than interior linemen on average, as they are more easily able to approach the quarterback on the edge of the pocket versus, for instance, a nose tackle attacking the line head on.

On average, STRAIN appears to increase for the first 0.5 seconds of a play, followed by a decline in the next second.
STRAIN then increases again until around 2.5-3 seconds after the snap before trending down towards the end of the play.
In context, this reflects the actions that a pass-rusher initially moves towards the quarterback, but is then stopped by the offensive line while the quarterback drops back.
When the quarterback stops dropping, the rusher closes the gap and increases STRAIN, before slowing down later on.

## Ranking the best pass-rushers

<!--

To determine the most effective pass rushers, we aggregate across all frames played and calculate the expected STRAIN (xSTRAIN) for each player, which is computed by simply averaging the STRAIN across all the frames of a play. The following tables show the top performing pass rushers (with at least 100 plays) rated by xSTRAIN for the first 8 weeks of the 2021 season.

The top interior rusher for the first 8 weeks of the 2021 season according to our xSTRAIN metric was Aaron Donald, almost universally considered the best interior rusher in the NFL currently with an xSTRAIN of 1.67. Through the first 8 weeks of the 2021 season, Donald had an impressive 5 sacks and 12 quarterback hits. Donald Was followed by two Raiders: Solomon Thomas and Quinton Jefferson with xSTRAIN values of 1.51 and 1.46, respectively. Rounding out the top 5 were Chris Jones with an xSTRAIN of 1.42 and DeForest Buckner with an xSTRAIN of 1.26.

As for edge rushers, Rashan Gary with an xSTRAIN value of 2.82 is the top of our list for the first 8 weeks of the 2021 season. Over this time period, Gary had 3.5 sacks and 15 quarterback hits. Second on our list of edge rushers is Leonard Floyd with an xSTRAIN value of 2.80 who had 6.5 sacks and 9 quarterback hits over the time period observed. Rounding out the top 5 are Justin Houston (2.78), Myles Garrett (2.75), and Von Miller (2.75). The sixth player on our rankings of edge rushers over the first 8 weeks on the 2021 season is the 2021 Defensive Player of the Year, T.J. Watt (2.71) who had 11.5 sacks and 18 quarterback hits in those first 8 weeks. Notably their is considerable overlap between our top 20 list over the first 8 weeks of the 2021 season and the Pro Football Focus edge rusher rankings prior to the 2022 season.

-->

To determine the most effective pass-rushers, we compute $\overline{\text{STRAIN}}$, which represents the average STRAIN across all frames played for each player.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass-rushers (NT's and DT's) separately from edge rushers (OLB's and DE's).
The following tables show the top performing pass-rushers (with at least 100 plays) rated by $\overline{\text{STRAIN}}$ for the first 8 weeks of the 2021 season.

Our results show many true positives with no surprises.
Myles Garrett and TJ Watt are known as top-tier edge rushers, while Aaron Donald is undoubtedly the best interior defender in football.
Rashan Gary, a young rusher with huge potential, ranks first among all pass-rushers according to our metric.
Our leaderboards largely match the rankings of experts in the field (for instance, PFF's [edge](https://www.pff.com/news/nfl-2022-edge-rusher-rankings-tiers) and [interior](https://www.pff.com/news/nfl-interior-defensive-line-rankings-tiers-2022) rusher rankings after the 2021 season), lending credibility to our proposed metric as a measure of pass-rushing effectiveness.

# Discussion

<!--

We have proposed a simple, effective, and easily interpretable metric for evaluating the performance of pass rushers in the NFL that we call STRAIN. Higher values of STRAIN indicate two important aspects of a pass rush: 1) the distance between the pass rusher and the quarterback is small and 2) the rate at which the pass rusher is moving towards the quarterback is high.

One of the major advantages of STRAIN over existing measures is that it gives a measure of effectiveness of a pass rush for every single play at a much more granular level than simply the play resulted in a sack or not or the play was a "success" or not. It is important to have a quantitative measure of pass rushing effectiveness if one wants to objectively evaluate the performance of pass rushers. STRAIN is just such a measure that objectively measures the pressure that pass rushers put on a quarterback.

In this report, we simply propose the new metric STRAIN; However, there are innumerable future projects that could be based on STRAIN. For example, STRAIN could be used to collectively evaluate the performance of a pass rushing unit by looking at the total amount of STRAIN during a play, which could, in turn, be used to evaluate the effectiveness of different types of blitzes. In addtion, STRAIN could be included as a covariate in models trying to predict quantities such as passing yards or completion probability. Notably, I believe there is very interesting work that could be done with functional regression (i.e. treating the STRAIN curve over time as a function) in modeling how STRAIN affects offensive production though is non-trivial as the length in time of plays varies from one play to another, and we leave this as future work.

In addtion, while we focused solely on evaluating the performance of pass rushers, this concept can be easily extended to the evaluation of pass blockers as a unit by simply looking at a quantity such as the total amount of STRAIN on a play or the maximum amount of STRAIN per frame aggregated across the entire play. Further, STRAIN could be extended to individual offensive pass blockers provided that a method of matching blockers to rushers was devised.

While STRAIN is a big step forward, there are still some notable limitations to this metric. For instance, STRAIN, as currently devised, cannot distinguish between a pass rusher who has to face a single blocker versus a double team. This problem could be addressed, at least somewhat, by devising a method of matching pass rushers to blockers as previously mentioned. Another limitation of this metric occurs when a quarterback scrambles out of the pocket. If you are a pass rusher on the left side and the quarterback scrambles out to the left, the STRAIN value for that pass rusher can drop considerably as the distance between the quarterback and the pass rusher gets larger, which the pass rushers shouldn't necessarily be negatively affected by in a measure of pass rushing effectiveness.

Finally, we believe there are other concepts in materials science that could be applied to evaluating pass rushing and pass blocking. One potential idea is a quantity called stress, which measures force over an area. Recall from intro physics that force is mass times acceleration, so this quantity would take into account the physical size of a pass rusher in the computation of force. You could then somehow divide this force over the "area" of the pocket formed by pass blockers to compute stress. There is also a quantity measuring the ratio of stress and strain called the elastic modulus which measures "an object's or substance's resistance to being deformed elastically". It is easy to imagine the pocket as an object and evaluating the succes of the pocket based on it's "resistance to being deformed".

-->

We have proposed STRAIN, a simple and interpretable statistic for evaluating pass-rushers, with higher STRAIN corresponds to greater pass-rushing ability.
Compared to existing metrics such as [ESPN's Pass Rusher Win Rate (PRWR)](https://www.espn.com/nfl/story/_/id/24892208), STRAIN measures pass-rush effectiveness for every play continuously over time, which is at a much more granular level than considering whether the play resulted in a sack or converting continuous data to a binary win/loss.

As for future directions, STRAIN could be utilized to collectively evaluate the performance of a pass-rushing unit by considering the total amount of STRAIN during a play, which could, in turn, be used to examine the effectiveness of different types of blitzes.
Moreover, while we focused solely on pass-rushers, STRAIN can be easily extended to the assessment of pass-blockers *as a unit* by looking at quantities such as the total STRAIN or the maximum STRAIN per frame aggregated across the entire play.
It is also possible to apply STRAIN to assess individual offensive pass-blockers, provided that a method of matching blockers to rushers was developed.

Our metric, however, is subject to several limitations.
As currently devised, STRAIN cannot distinguish between a pass-rusher who has to face a single blocker versus a double team.
This problem could be addressed, at least somewhat, by having a method of matching pass-rushers to blockers as previously mentioned.
Additionally, there are a few situations such as quarterback out-of-pocket scrambling where STRAIN does not appear to be useful.
If a pass-rusher is on the left side and the quarterback scrambles out to the left, the STRAIN for that pass-rusher can drop considerably as the distance between the quarterback and pass-rusher increases, which the defender should not necessarily be negatively affected by in a measure of pass-rushing effectiveness.