### 2023-03-10 paper/02-methods.Rmd
separate files for each section

---
output: rticles::rss_article
bibliography: bibliography.bib
---

# Methods {#sec:methods}

## Data

2023 bdb [@Howard2023NFL]

first 8 weeks 2021 season

tracking: locations of players and ball, 10 fps,

pre-processing standardize

PFF scouting: pass rush role

sample restriction: remove plays w multiple QB

frames between snap and QB outcome (pass forward, sack)

sample size

## Strain rate in materials science

math

## Application to pass-rushing in football

analogy