### 2023-05-11 paper/00-main.Rmd
more abstract edits

---
title: |
  | \Large Bearing the STRAIN: Analyzing Pass Rush in American Football with Player Tracking Data

# Deforming the line:
# Straining the line:
# Unleashing the Rush
# straining for the takedown
# straining for success
# straining forward
# straining for pressure
# Unleashing the Strain
# Strain on the Gridiron
# from strain to gain
# Bearing the Strain:
# Straining the trenches:

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Quang Nguyen
  # thanks: Corresponding author's email
  affiliation: |
    | Department of Statistics & Data Science
    | Carnegie Mellon University
- name: Ronald Yurko
  affiliation: |
    | Department of Statistics & Data Science
    | Carnegie Mellon University
- name: Gregory J. Matthews
  affiliation: |
    | Department of Mathematics and Statistics
    | Loyola University Chicago

keywords:
- tracking data
- football
- pass rush
- statistics in sports

abstract: |
  In American football, a pass rush is an attempt by the defensive team to disrupt the offense and prevent the quarterback (QB) from completing a pass. Existing metrics for assessing pass rush are either discrete-time quantities or based on subjective judgement. Using player tracking data, we propose STRAIN, a novel metric for evaluating pass rushers in the National Football League (NFL) at the continuous-time within-play level. Inspired by the concept of strain rate in materials science, STRAIN is a simple and interpretable means for measuring defensive pressure in football. It is a directly-observed statistic as a function of two features: the distance between the pass rusher and QB, and the rate at which this distance is being reduced. Our metric possesses great predictability of pressure and stability over time. We use a multilevel model to isolate the contribution of every pass rusher in a play to the average STRAIN while controlling for their pass blocker matchup, the QB, and play-level information. We apply our approach to NFL data and present results for the first 8 weeks of the 2021 season. In particular, we provide comparisons of STRAIN for different defensive positions and play outcomes, and rankings of the NFL's best pass rushers according to our metric.

bibliography: bibliography.bib
biblio-style: chicago
output: rticles::asa_article
header-includes:
  # \usepackage{mathptmx}
  # \renewcommand{\arraystretch}{0.8}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.pos = "p",
  cache = TRUE
)
library(here)
```

\newpage

```{r child = "01-introduction.Rmd"}
```

```{r child = "02-data.Rmd"}
```

```{r child = "03-methods.Rmd"}
```

```{r child = "04-results.Rmd"}
```

```{r child = "05-discussion.Rmd"}
```

# Acknowledgements {-}

Big data bowl, Mike

\bigskip
\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

**GitHub repository:** https://github.com/getstrained/intro-strain.

**NFL Big Data Bowl 2023 Data:**: https://www.kaggle.com/competitions/nfl-big-data-bowl-2023/data.

# References