### 2023-07-05 paper/supplemental.Rmd

---
title: |
  | Supplementary materials for
  | "Here Comes the STRAIN: Analyzing Defensive Pass Rush in American Football with Player Tracking Data"
blinded: 0
bibliography: bibliography.bib
output: rticles::asa_article
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.pos = "p"
)
library(here)
library(tidyverse)
```